I/O simulation model with Linux page cache, integration and evaluation in SimGrid framework

### 2020-10-19 paper_io_sim.tex
Update paper

>>>>

		We use the disk and memory benchmark results and CPU time measured from 
		the cluster to configure the simulated platform. The benchmark results of 
		bandwidths and configured values in the simulators are shown in \ref{table:benchmark}.

		\begin{table}
		\centering
		\begin{tabular}{|l|c|c|c|}
		\hline
			Bandwidths (MBps)  & Cluster (real) & Python simulator & WRENCH simulator\\
		\hline
			Memory read  & 6860	& 4812	 & 4812\\
			Memory write & 2764	& 4812 & 4812\\
			Disk read & ~510 & 465 & 465\\
			Disk write & ~420 & 465	 & 465\\
		\hline
		\end{tabular}
		\caption{Bandwidth benchmark results and simulator configurations}
		\label{table:benchmark}
		\end{table}

		As read and write bandwidths are symmetrical in SimGrid, we take the average values 
		to configure our simulators.

		\subsection{Single-threaded experiment}

			Memory behaviors of real pipeline execution and simulators are recorded 
			and the memory profiling results with 20GB and 100GB input files 
			are illustrated in Figure~\ref{fig:single_memprof_20gb} and Figure~\ref{fig:single_memprof_100gb}. 
			The bars in the figures present activities in tasks (read, compute and write), 
			while the lines describe memory status throughout run time of the pipeline. 
			Because the results with 20Gb, 50GB, 100GB are analogous, we only include 
			the results with 20GB and 100GB in this paper.

			As shown in the figures, the makespans of the simulated pipeline are 
			close to the makespan of the real execution with similar patterns of I/O time 
			real execution and simulation. 
			But more importantly, with all input sizes, similar trends can also be found 
			in the lines describing memory status in simulators and the real execution. 
			The akin movements of amount of cache used (purple lines) in reality 
			and simulation show that the use of page cache is well reproduced by the simulators. 
			The amount of dirty data ( red lines) also follow similar trends, 
			which means not only the amount dirty data, but also the flushing and 
			periodical flushing mechanisms are accurately modeled in our simulators. 
			This can be seen in in the increase of the amount of dirty data during the writes 
			and the decrease during the reads and computations. 
			According to the results achieved with all input sizes in the experiment, 
			we can observe identical trends in reality and simulation with 20GB, 
			50GB and 75GB input files. However, with 100GB input file, 
			the third read is not correctly simulated, leading to longer simulated time 
			than in real execution. This will be explained later in this section.

			The I/O time of tasks in simulators and real execution is also tracked and compared. 
			Figure~\ref{fig:single_error_20} and Figure~\ref{fig:single_error_100} show 
			the relative error of simulated I/O time of individual I/O activities. 
			Our pipeline consists of 3 tasks, and each task reads an input file, 
			performs some computation, and writes output file. 
			Thus, there are 6 sequential I/O activities, \textit{read1} and \textit{write1} 
			from the first task, \textit{read2} and \textit{write2} the second task, 
			and \textit{read3}, \textit{write3} from the last task.

			In general, with all input sizes in the experiment, the simulation errors 
			of simulators with page cache model (Python and WRENCH with page cache simulators) 
			are remarkably lower than the result with original WRENCH . 
			The reason is, in original WRENCH, because page cache is not simulated, 
			data is only read/written with disk bandwidth, which is much slower 
			than memory bandwidths, leading to simulated I/O time longer than in reality.
			Especially, the simulation time errors of \textit{read3} in the simulators 
			with page cache are just under the error in original WRENCH.

			\begin{figure*}
			\begin{subfigure}{\columnwidth}
				\centering
   				\includegraphics[width=\columnwidth]{result/single/figures/single_comp_sep_20gb.pdf}
   				\caption{Memory profiling with 20GB input file}	\label{fig:single_memprof_20gb}
			\end{subfigure}	
			\begin{subfigure}{\columnwidth}
				\centering
   				\includegraphics[width=\columnwidth]{result/single/figures/single_comp_sep_100gb.pdf}
   				\caption{Memory profiling with 100GB input file}	\label{fig:single_memprof_100gb}
			\end{subfigure}	
			\begin{subfigure}{\columnwidth}
				\centering
   				\includegraphics[width=\columnwidth]{result/single/figures/error_20.pdf}
   				\caption{Simulation error with 20GB input file}	\label{fig:single_error_20}
			\end{subfigure}		
			\begin{subfigure}{\columnwidth}
				\centering
   				\includegraphics[width=\columnwidth]{result/single/figures/error_100.pdf}
   				\caption{Simulation error with 100GB input file}	\label{fig:single_error_100}
			\end{subfigure}	
			\caption{Simulation results with different input file sizes}
			\label{fig:single_results}
			\end{figure*}

			\begin{figure*}
			\begin{subfigure}{\columnwidth}
				\centering
   				\includegraphics[width=\columnwidth]{result/single/figures/fincore_real.pdf}
   				\caption{Real pipeline execution}	
   				\label{fig:fincore_real}
			\end{subfigure}	
			\begin{subfigure}{\columnwidth}
				\centering
   				\includegraphics[width=\columnwidth]{result/single/figures/fincore_wrench.pdf}
   				\caption{WRENCH with page cache simulator}	
   				\label{fig:fincore_wrench}
			\end{subfigure}	
			\caption{Amount of file data in cache after each I/O activity with 100GB of input}
			\label{fig:fincore}
			\end{figure*}

			This longer read can be explained with the amount of cached data of files 
			at the beginning of the read in reality and simulators shown in Figure~\ref{fig:fincore}. 			
			In this figure, \textit{file$<$i$>$} is the input of \textit{read$<$i$>$} and 
			output of \textit{write$<$i-1$>$}. 
			In the real execution, after \textit{write2}, which writes \textit{file3}, 
			all data of \textit{file3} is in cache (Figure~\ref{fig:fincore_real}). 
			Thus, in \textit{read3}, \textit{file3} is read with memory bandwidth. 
			But in our simulator, after \textit{write2}, only a part of \textit{file3} 
			remains in cache (Figure~\ref{fig:fincore_wrench}). 
			The reason is while \textit{file3}  is being written, the page cache is saturated, 
			eviction is then evoked to evict data from page cache, make space 
			available for \textit{file3}. 
			However, because the evicted amount is less than required, 
			some \textit{file3} data is evicted after being written, make \textit{file3} 
			partially cached. 
			As a result, in \textit{read3}, only a part of \textit{file3} is read from cache 
			with memory bandwidth, a part of it is read from disk with disk read bandwidth, 
			which is many times slower, leading to longer read time.

			In summary, in this single-threaded experiment, our model can accurately 
			simulate I/O time as well as the internal memory behavior. 
			This can help confirm that our model is on a right track and 
			it can be further improved to achieve higher accuracy.

		\subsection{Multi-threaded experiment}

			\begin{figure*}
			\begin{subfigure}{0.33\linewidth}
				\centering
   				\includegraphics[width=\linewidth]{result/multi/figures/makespan.pdf}
   				\caption{Total makespan}	
   				\label{fig:multi_makespan}
			\end{subfigure}	
			\begin{subfigure}{0.33\linewidth}
				\centering
   				\includegraphics[width=\linewidth]{result/multi/figures/readtime.pdf}
   				\caption{Average read time}	
   				\label{fig:multi_readtime}
			\end{subfigure}	
			\begin{subfigure}{0.33\linewidth}
				\centering
   				\includegraphics[width=\linewidth]{result/multi/figures/writetime.pdf}
   				\caption{Average write time}	
   				\label{fig:multi_writetime}
			\end{subfigure}		
			\caption{Total makespan and I/O time with different number of concurrent pipeline}
			\label{fig:multi_result}
			\end{figure*}

			In this experiment, the results is analyzed in terms of the total makespan  
			and average I/O time concurrent pipelines.
			The results from real execution and simulation are presented in Figure~\ref{fig:multi_result}.

			The Figure~\ref{fig:multi_makespan} shows that as the number of pipelines 
			increases, the total makespan raises accordingly because the amount of data 
			increases but the memory and disk bandwidths are fixed and shared between pipelines. 
			In comparison with original WRENCH, the result from WRENCH with page cache are closer 
			to reality. 
			This proximity can also be seen in average read time shown in 
			Figure~\ref{fig:multi_readtime}. 
			Especially, when it comes to the average write time 
			(shown in Figure~\ref{fig:multi_writetime}), 
			although the trend in reality is more complex, our WRENCH with page cache 
			simulator can still capture it and share the same pattern with 
			the real execution, when they gradually increase before surging 
			with the same slope after the number of pipelines reaches 10. 
			This can be explained that with less than 10 concurrent pipelines, 
			the page cache is not saturated, so all files can be written entirely to 
			cache with memory bandwidth in a short time. 
			After the page cache is saturated with dirty data (at around 
			10 concurrent pipelines in this experiment), this dirty data needs 
			to be flushed in order to make space available for writing new data to cache. 
			This dirty data is flushed to disk with disk bandwidth, which is much  
			slower than memory bandwidth. 
			The more pipelines we have, the more data needs to be written, the more data 
			needs to be flushed, leading to a sharper increase in the average write time.

			To conclude, the results from WRENCH with page cache simulator show that 
			our model can also simulate I/O with page cache accurately in a 
			multi-threaded application.

		\subsection{Remote storage}

		\subsection{Real application}
